---
title: "DeepNeuralNetworks4R"
author: "Óscar González-Velasco"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

================
Implementation of _Deep Neural Networks_ in R programing language.
----------------

Regression algorithm for Omic data prediction in brain transcriptomics (although as a regression model, it can be applied to **any** problem with a dependent continuous variable).

We will use *_a set of transcriptomic data from human brain samples_* included on the package as an example of a *_regression_* model using *deep neural networks* to predict the biological age:

```{r }
# We load the Deep Neural Network package:
library(DeepNeuralNetworks4R)

```

We will try to predict the age of the individuals based on the gene expression of 1078 genes selected because of its implications on brain aging on cortex region (Oscar González-Velasco, et al., BBA - Gene Regulatory Mechanisms, https://doi.org/10.1016/j.bbagrm.2020.194491).

```{r }
# We inspectionate the data included within the package:
training.data[1:5,1:5]

# We will select the first 3 genes (the most significant genes linked with aging) to build 3 additional
# data matrix using each of these genes as the centroid:
zscore.targets <- as.list(rownames(training.data))[1:3]

# Print the 3 first genes
zscore.targets
```
Here we can find an except for the training dataset, *notice* that *response variables* correspond to *_rows_*, meanwhile *samples* correspond to *_columns_* .

## Training the regression model

First, we proceed to create the deep neural network model:

```{r, message = FALSE}
model <- deepNeuralNetwork.build(
            x=1:(nrow(training.data)-1),
            y=nrow(training.data),
            outputNeurons = 1,
            HidenLayerNeurons = c(20,20,20,10,10,10,10,10,10,10,10,10,10,10,10,10,10,10),
            traindata=training.data,
            drawDNN = 0,
            standarization = zscore.targets)
```

*_x_* will specify the index positions of our explanatory variables on the matrix _training.data_
*_y_* will specify the index positions of our observed variable on the matrix _training.data_, here it will correspond with the age.

*_HidenLayerNeurons_* will specify the number of neurons that each layer will have. The number of neurons on the very first layer will be the number of variables that we will use to create the regression model (deepNeuralNetwork.build calculate this automatically based on the x parameter).

*_deepNeuralNetwork.build_* will create an object of class _DeepNNModel_ that will store all the information about the dnn model.

And now we train the deep neural network using the following code:

```{r, message=FALSE}
# 3. train model
timeNN <- system.time(
  model.trained <- deepNeuralNetwork.training(
                        x=1:(nrow(training.data)-1),
                        y=nrow(training.data),
                        model = model, #ddn.model.in.use,
                        traindata=training.data,
                        testdata=test.data,
                        iterations  = 1000,
                        lr = 0.001,
                        reg = 0.001,
                        display=1000,
                        maxError = 0.1,
                        standarization = zscore.targets))

```

## Testing the results

Once we have the model, we will make use of *_deepNeuralNetwork.predict_* function to predict a variable based on the trained regression model:

```{r message = FALSE,fig.dim = c(10, 10)}

age.prediction <- deepNeuralNetwork.predict(model.trained = model.trained,
                                            data = test.data[-nrow(test.data),],
                                            standarization = zscore.targets)


mplot_lineal(observed = test.data[nrow(test.data),],
             predicted = age.prediction,
             title = "Biological age prediction using DNN regression from human brain data",
             x.lim = c(1,7),
             y.lim = c(1,7),
             x.lab="chronological age (observed)",y.lab = "predicted bio-age (predicted)")

```

## Using GPU for large datasets

The DeepNN algorithm has been optimized to be executed on a GPU card using R's matrix/vector arithmetic expressions.

The following example shows how to run the DNN on a GPU using CUDA with R's nvblas config file:

```{bash, eval=FALSE}
sudo env LD_PRELOAD=/PATH/TO/CUDA/NVBLAS/libnvblas.so NVBLAS_CONFIG_FILE=/PATH/TO/NVBLAS.CONFIG.FILE/nvblas.conf R CMD BATCH ./regression.deepNN.GPU.r /dev/tty
```
